---
title: "Biostats 597 HW1"
author: Bing Miu
output:
  pdf_document:
    latex_engine: xelatex
  html_document:
    df_print: paged
header-includes:
  - \usepackage{color}
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE)
library(tidyverse)
library(gapminder)
```

A look at the data:
```{r echo=FALSE, warning=FALSE, message=FALSE}
gapminder
```

# HW exercises: 
use ggplot to make the graphs requested in (1) and (2). Throughout, avoid repetition of code.
Goal: gain experience with ggplot and reading help files/using google searches to get help on (arguments of) R functions. 

1. make one scatter plot using ggplot with
- Life expectancy on the y-axis, label axis as "Life expectancy"
- GDP per capita on the x-axis, use a log-transform of the x-axis while still displaying the (unlogged) GDP values 
- shapes to indicate the continent, add a legend
- color to indicate the year, do NOT add a legend for that

```{r}
ggplot(data=gapminder, mapping = aes(x = gdpPercap, y=lifeExp)) +
  labs(title= "Life expectancy at birth VS GDP per capital", x = "GDP per Capita (PPP)", y = "Life expectancy (years)") +
  scale_x_continuous(trans = "log10") +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_point(mapping = aes(color=year, shape=continent)) +
  scale_color_continuous(guide= 'none')
```



2. add two geom_smooth layers (a) and (b) to the plot you made in (1) where
(a) is fitted to the entire dataset and uses default settings for the smoother used and plotting settings


```{r}
ggplot(data=gapminder, mapping = aes(x = gdpPercap, y=lifeExp)) +
  labs(title= "Life expectancy at birth VS GDP per capital", x = "GDP per Capita (PPP)", y = "Life expectancy (years)") +
  scale_x_continuous(trans = "log10") +
  theme(plot.title = element_text(hjust = 0.5))+
  geom_point(mapping = aes(color=year, shape=continent)) +
  scale_color_continuous(guide= 'none')+
  geom_smooth()
```

(b) is fitted to data in Africa only, provides a 99% confidence interval, is plotted in red (both point estimates as well as confidence interval), and is transparent (so if it overlaps with (a), (a) is still visible). 
Hint: to fit to data in Africa only, consider using "data = filter(gapminder, continent == "Africa")"

```{r}
ggplot(data=gapminder, mapping = aes(x = gdpPercap, y=lifeExp)) +
  labs(title= "Life expectancy at birth VS GDP per capital", x = "GDP per Capita (PPP)", y = "Life expectancy (years)") +
  scale_x_continuous(trans = "log10") +
  theme(plot.title = element_text(hjust = 0.5))+ 
  scale_color_continuous(guide= 'none') +
  geom_point(mapping = aes(color=year, shape=continent)) +
  geom_smooth() +
    geom_smooth(data = filter(gapminder, continent== "Africa"), color="red", fill ='red', level = 0.99, alpha = 0.2)
```

\color{red}Red = fitted to data in Africa only    
\color{blue}Blue = fitted to data in all continents

